---
title: "An exploration of fish communities in California MPAs"
format: html
execute:
  echo: false
  warning: false
  message: false
embed-resources: true
---

```{r}
library(mosaic)
library(tidyverse)
library(ggplot2)
library(ggridges)
drop <- read_csv("raw-data/drop.csv", col_types = list(Drop_ID = col_character()))
points <- read_csv("raw-data/points.csv", col_types = list(OpCode = col_character()))
lengths <- read_csv("raw-data/lengths.csv", col_types = list(OpCode = col_character()))

```


### Data Description

These datasets are from a summer of data collected in 2023, with the purpose of assessing the possible impacts of marine protected areas (MPAs) on the fish communities they protect. That summer, the Lindholm lab and I completed four week long data collection trips at four different MPAs along California's central and north coast. We deployed stereo camera systems for 45 minutes each, then took that video back to the lab and spent the fall collecting fish data from it. Stereo video specifically uses two cameras to triangulate objects where the frames overlap, allowing us to measure fish down to the millimeter without ever touching them. These data were eventually processed and submitted to inform MPA management.

Here is the data: https://docs.google.com/spreadsheets/d/1KVd_7IsOHM9Xe1PDIf47Q4LnQt4J6QFjS5UL1i_GzpA/edit?gid=0#gid=0 

*Sheets*:

- Drop_data - information on each camera system deployment <br>
- 3D_points - Each individual fish counted <br>
- Lengths - The precise measurement of each fish able to be measured

*Variables of interest*:

- OpCode OR Drop_ID - sample (drop) identifier <br>
- Date - YYYYMMDD, when the video was taken <br>
- Location - the marine protected area that was sampled <br>
- Designation - whether the video was taken inside the MPA or outside (REF) <br>
- 3D Points - number of fish counted in drop <br>
- 3D Lengths - number of fish measured in drop <br>
- Depth (m) - depth of drop in meters <br>
- Simpson_div - Gini-Simpson diversity index <br>
- Shannon_div - Shannon-Wiener diversity index <br>
- Common_name - common name of each species of fish

### Data Cleaning

For the first visualization, I started with the 3D_points data and pulled out just the columns of OpCode and Common_name. Then, grouping by OpCode, I counted the number of unique values for Common_name. This is calculating the number of different species seen per sample, or species richness. I then take this information and add it to the Drop_data sheet. I then make minor changes such as renaming the species richness variable and ordering the locations from south to north before making the visualization. <br><br> For the second visualization, I used the lengths sheet, and pulled out columns Date, Location, Designation, Length (mm), OpCode, and Common_name. I then removed observations of Northern Anchovies. This was a pretty extreme outlier, where in one video, a bait ball of thousands of anchovies swam in view of the cameras. This was severely skewing the data, and only happened in one out of over 200 drops, so it was removed. Then I converted the lengths from millimeters to centimeters and ordered the locations from south to north before making the seconf visualization.


```{r}

points_clean <- points |> 
  select(OpCode, Common_name) |> 
  group_by(OpCode) |> 
  summarize(n_distinct(Common_name)) |> 
  left_join(drop, by = c("OpCode" = "Drop_ID")) |> 
  rename(Spp_richness = 'n_distinct(Common_name)') |> 
  mutate(Location = factor(Location, levels = c("Point_Buchon", "Point_Lobos", "Ano_Nuevo", "Montara")))

write_csv(points_clean, file = "clean-data/points-clean.csv")

```

```{r}
lengths_clean <- lengths |> 
  select(Date, Location, Designation, `Length (mm)`, OpCode, Common_name) |>
  filter(Common_name != "Northern_Anchovy") |> 
  mutate(length_cm = `Length (mm)`/10) |> 
  mutate(Location = factor(Location, levels = c("Point_Buchon", "Point_Lobos", "Ano_Nuevo", "Montara"))) 

write_csv(lengths_clean, file = "clean-data/lengths-clean.csv")
```

### Figure 1

```{r}
points_clean|> 
  ggplot(aes(x = Designation, 
             y = Spp_richness, 
             color = Location, 
             fill = Designation)) +
  geom_boxplot() +
  facet_grid(~ Location)+
  labs(x = "Species Richness", 
       y = "MPA Designation", 
       title = "Species richness is generally higher in MPAs than reference sites",
       alt = "There are four sets of two boxplots, from locations Point Buchon, Point Lobos, Ano Nuevo, and Montara. Each site has a boxplot for within the MPA and outside the MPA. The y-xis is species richness, which ranges from 0 to 12.5. The species richness appears higher in every MPA except Ano Nuevo.") +
  scale_color_viridis_d(end = 0.7) +
  scale_fill_manual(values = c("lightgray", "darkgray")) +
  theme_minimal() +
  guides(fill = "none", 
         color = "none")
```

### Figure 2

```{r}
lengths_clean |>   
  ggplot() +
  geom_density_ridges(aes(x = length_cm, 
                          y = Location, 
                          color = Location), 
                      jittered_points = TRUE, 
                      alpha = 0.5) +
  labs(y = "Marine Protected Area Site",
       x = "Length (cm)",
       title = "Distributions of fish size by site",
       alt = "There are four density ridges, one for each site mentioned previously. The x axis is the size of fish in centimeters. The density ridges are filled with points each representing an individual fish. The ridges get higher based on how many points are in that size range. Ano Nuevo, Point Lobos, and Point Buchon had seemingly more big fish than Montara.") +
  scale_color_viridis_d(end = 0.7) +
  guides(color = "none") +
  theme_minimal()
  
```
